The kinematics of the Local Group in a cosmological context

### 2013-03-01 paper/reply_referee.txt

Dear Editor,

We have appreciated the positive referee's comments and the suggestions
for improvement in the paper.

In what follows below the dashes, the referees comments are indented
and our responses are not commented.

Jaime E. Forero-Romero (on behalf of the authors)

----

	First, it would be useful if the authors were more explicit about how
	they are defining "consistency". Reading the text fairly carefully, I
	think that I was able to discern that you are using 1-sigma
	constraints on v_tan to define consistency, but this needs to be
	stated very clearly. For example, what is the quoted c.l. for the HST
	limit on V_tan listed in Table 1? This should be stated clearly in the
	table itself. In the caption of Table 3, again describe exactly how
	you are defining consistency. It would be useful to know how these
	numbers change as you very the definition (1-sigma, 2, 3?). 
	I discern from a footnote that the contours drawn in Figure 1 are
	1-sigma contours. Please state this clearly in the caption
	and/or in the text that describes the figure.

We have now included explicit mention of the 1-sigma criteria in Table
1, Table 3 and the caption of Figure 1.

   	 Second, the authors make a good case pairs that have Vr/Vtan values
	 similar to those observed are somewhat rare in the simulations and
	 suggest that this might possibly imply that the LG is not typical. It
	 should be possible to ask some basic questions about the small # of
	 systems that appear consistent with the data: does it matter? Do they
	 have outlying halo spins, concentrations, or formation times? What
	 about their larger-scale environments?

The answers to these questions are being clarified in a paper under
preparation by one of the coauthors (S. Bustamante).

The largest difficulty to study the pairs consistent with Vr/Vtan is
their low number. There are only 8 of them for the general sample and
_none_ in the reduced sample. This makes any statistical claim hard to
obtain. However we have found that these pairs prefer special large
scale environments defined by the tidal and velocity shear tensor
fields, namely they live in filaments. We are also investigating their
alignment with respect to the directions along this filaments. We will
include in that study the referee sugestions to study the
concentration and spin parameter.

We consider that this low number matters to the extent that it shows
how unlikely is to have Local Group like kinematic configurations in
LCDM, which might be relevant to explain other features in the matter
density field such as the special alignment of satellites in the Milky
Way and Andromeda or might hint towards a constraint on the expected
MW assembly history in LCDM.

### 2013-03-01 paper/reply_referee.txt

Dear Editor,

We have appreciated the positive referee's comments and the suggestions
for improvement in the paper.

In what follows below the dashes, the referees comments are indented
and our responses are not indented.

The refereee raised concerns on Figures that look cut off. It seems to
be a technical problem on the side of the ApJLetter submissions
software. The raw eps figures and the pdf we produce (latex, dvips,
ps2pdf) look complete (see:
https://github.com/forero/LG_Kinematics/blob/master/paper/ms.pdf ,
click on 'View Raw').  We kindly as the Editor to pay attention to
this issue and inform us if we should make any further change to our
figures.

With best regards,

@@

Jaime E. Forero-Romero (on behalf of the authors)

----

	First, it would be useful if the authors were more explicit about how
	they are defining "consistency". Reading the text fairly carefully, I
	think that I was able to discern that you are using 1-sigma
	constraints on v_tan to define consistency, but this needs to be
	stated very clearly. For example, what is the quoted c.l. for the HST
	limit on V_tan listed in Table 1? This should be stated clearly in the
	table itself. In the caption of Table 3, again describe exactly how
	you are defining consistency. It would be useful to know how these
	numbers change as you very the definition (1-sigma, 2, 3?). 
	I discern from a footnote that the contours drawn in Figure 1 are
	1-sigma contours. Please state this clearly in the caption
	and/or in the text that describes the figure.

We have now included explicit mention of the 1-sigma criteria in Table
1, Table 3 and the caption of Figure 1. There is also a new paragraph in Section 3.1 dicussing the results for 3-sigma uncertainties.

@@

   	 Second, the authors make a good case pairs that have Vr/Vtan values
	 similar to those observed are somewhat rare in the simulations and
	 suggest that this might possibly imply that the LG is not typical. It
	 should be possible to ask some basic questions about the small # of
	 systems that appear consistent with the data: does it matter? Do they
	 have outlying halo spins, concentrations, or formation times? What
	 about their larger-scale environments?

The answers to these questions are being clarified in a paper under
preparation by one of the coauthors (S. Bustamante).

The largest difficulty to study the pairs consistent with Vr/Vtan is
their low number. There are only 8 of them for the general sample and
_none_ in the reduced sample. This makes any statistical claim hard to
obtain. However, we have found that these pairs prefer special large
scale environments defined by the tidal and velocity shear tensor
fields, namely they live in filaments. We are also investigating their
alignment with respect to the directions along this filaments. We will
include in that study the referee sugestions to study the
concentration and spin parameter.

@@

We consider that this low number matters to the extent that it shows
how unlikely is to have Local Group like kinematic configurations in
LCDM, which might be relevant to explain other features in the matter
density field such as the special alignment of satellites in the Milky
Way and Andromeda or might hint towards a constraint on the expected
MW assembly history in LCDM.

   	    Third, while this is minor in terms of content, I noticed
	    a number of	grammatical issues. I strongly suggest a
	    re-read for type-os and wording.

We have fixed all the instances noted by the referee in the report.

Concerning the Figures that look cut off, it seems a technical problem
of the ApJLetter software. The raw eps figures and the pdf we produce
(latex, dvips, ps2pdf) look complete (see:
https://github.com/forero/LG_Kinematics/blob/master/paper/ms.pdf ,
click on 'View Raw')

### 2013-03-01 paper/ms.tex
matches resubmitted version
updated version

\begin{abstract}
Recent observations constrained the tangential velocity of M31 with
respect to the Milky Way (MW) to be $v_{\rm M31,tan}<34.4$ \kms and
the radial velocity to be in the range $v_{\rm M31,rad}=-109\pm
4.4$\kms \citep{vanderMarel12}. In this study we use a large volume
high resolution N-body cosmological simulation (Bolshoi) together with three
constrained simulations to statistically study this kinematics in the
context of the $\Lambda$CDM. The comparison of the ensembles of
simulated pairs with the observed LG has been done with respect to the
observed radial and tangential velocities, the reduced orbital energy
($e_{\rm tot}$), angular momentum ($l_{\rm orb}$) and the
dimensionless spin parameter, $\lambda$. Our main results are: (i) the
preferred radial and tangential velocities for pairs in $\Lambda$CDM
are $v_{\rm r}=-80\pm 20$\kms, $v_{\rm   t}=50\pm 10$\kms, (ii) pairs
around that region are $3$ to $13$ times more common than pairs within
the observational values, (iii) $15\%$ to $24\%$ of LG-like pairs in
$\Lambda$CDM have energy and angular momentum consistent with
observations while (iv) $9\%$ to $13\%$ of pairs in the same sample
show similar values in the inferred dimensionless spin parameter. It
follows that within current observational uncertainties the
quasi-conserved quantities that characterize the orbit of
the LG, i.e. $e_{\rm tot}$, $l_{\rm   orb}$ and $\lambda$, do not
challenge the standard $\Lambda$CDM model, but the model is in tension
with regard to the actual values of the radial and tangential
velocities. This might hint to a problem of the $\Lambda$CDM model to
reproduce the observed LG.

@@

The Milky Way (MW) and Andromeda galaxy (M31) are the dominant
galaxies in the Local Group (LG). Astronomical observations of their
mass distribution impose constraints on the standard cosmological
model. The satellite overabundance problem \citep{Klypin99,Moore99},
tidal disruption features \citep{pandas09} and the disk dominated
morphology \citep{Kazantzidis2008} are examples on how LG studies
are linked to the cosmological context. Detailed studies on the
Magellanic Clouds dynamics and their possible link to M31 add to the
interest of understanding the details of the LG kinematics and
dynamics
\citep{Besla2007,Tollerud2011,Knebe2011,Fouquet2012,Teyssier2012}. However,
a general concern in the use of the LG as a tool for near-field
cosmology \citep{Freeman2002,Peebles2010} is how typical is the LG
regarding the properties of interest
\citep{Busha2011,Liu2011,ForeroRomero2011,Purcell2012}.

A new valuable piece of information in this issue is the recent
observational determination of the proper-motion measurements of M31,
which until recently had been out of reach \citep{vanderMarel12}.  The
reported measurements set an upper bound for the tangential velocity
of M31 with respect to the MW of $v_{\rm tan,M31}\leq 34.4$ km
s$^{-1}$. Together with the values of the relative radial velocity of
$v_{\rm rad,M31}=-109\pm 4.4$ \kms observations show that the relative
motion of the MW and Andromeda is consistent with a head-on
collision. With this information it is possible to quantify how common
is this kinematic configuration in a $\Lambda$CDM Universe.

This Letter presents such study. We use a large volume, high resolution
dark matter only N-body simulation in the concordance $\Lambda$CDM
cosmology to find a set of halo pairs with similar  characteristics as
inferred in the LG. We quantify these results in terms of the number
of pairs with given radial and tangential velocities in the
galactocentric rest frame. We also find the pairs that are consistent
with a head on collision in terms of the ratio of the radial to
tangential velocity $f_{\rm t}\equiv v_{\rm tan}/v_{\rm rad}<0.32$ and
present these results in terms of the reduced angular momentum,
mechanical energy and dimensionless spin parameter.

@@

This Letter is structured as follows. In the next section we present
the N-body simulations, the criteria we use to select LG-like halo
pairs. In Section \ref{sec:results} we present the results for the
dynamics in these pairs in terms of the tangential/radial velocities
and the orbital angular momentum/mechanical energy. In the same
section we summarize these dynamical results in terms of the
dimensionless spin parameter of the pairs. Finally, in the last
section we comment and conclude about the implications of these
results in the context of the $\Lambda$CDM model.

@@

\begin{table}
\caption{Summary of kinematic observational constraints.}
\begin{center}
\begin{tabular}{ccc}\hline
$v_{\rm M31,rad}$ &(\kms) & $-109.3\pm 4.4$\\
$v_{\rm M31,tan}$ &(\kms) & $<34.4$\\
$r_{\rm M31}$ &(kpc) & $770\pm 40$\\
${\bf r}_{\rm M31}$ & (kpc) &$(-378.9, 612.7, -283.1)$\\
$\sigma_{{\bf r}, {\rm M31}}$ & (kpc) &$(-18.9, 30.6, 14.5)$\\
${\bf v}_{\rm M31}$ & (\kms) & $(66.1, -76.3, 45.1)$\\
$\sigma_{{\bf v},{\rm M31}}$ & (\kms) &$(26.7, 19.0, 26.5)$\\
$M_{\rm 200, MW}$ & ($10^{12}\Msun$) & $1.6\pm0.5$ \\
$M_{\rm 200, M31}$ & ($10^{12}\Msun$) & $1.6\pm0.5$ \\
$M_{\rm 200,MW} + M_{\rm 200, M31}$ & ($10^{12}\Msun$) & $3.14\pm 0.58$\\
$\log_{10}\lambda$& & $-1.72\pm 0.07$ \\\hline
\end{tabular}\\
\end{center}
\vspace{1mm}
Notes:\\
\begin{enumerate}
\item The kinematic
  properties for M31 are reported in the galactocentric restframe 
  \citep{vanderMarel12}.

\item Values in parenthesis correspond to vector
  components. $\sigma_{\bf x}$ represents the uncertainty on the
  components of vector ${\bf x}$. 
\item The values for the individual halo
masses are consistent with the priors used by \cite{vanderMarel12}.

\item The observational uncertainties in the position vector
correspond to a $5\%$ in each component consistent with the
uncertainties in the distance \citep[see references
  in][]{vanderMarel08}.

@@

\begin{figure*}
\begin{center}
\includegraphics[keepaspectratio=true,width=0.46\textwidth]{fig_1a.eps}
\includegraphics[keepaspectratio=true,width=0.46\textwidth]{fig_1b.eps}
\includegraphics[keepaspectratio=true,width=0.46\textwidth]{fig_1c.eps}
\includegraphics[keepaspectratio=true,width=0.46\textwidth]{fig_1d.eps}
\caption{Upper panels: 2D Histograms of the radial and tangential
  velocities for LG-like halo pairs in the Bolshoi simulation.   Lower
  panels: 2D Histograms of the orbital angular momentum ($l_{\rm orb}$)
  and mechanical energy ($e_{\rm tot}$) per unit of reduced mass
  calculated considering the halos as point masses. The bar indicates
  the number of pairs in each cell. The locii for the high
  density regions in the $v_{\rm r}-v_{\rm t}$ plane are listed in
  Table \ref{table:2}. 
  The half ellipse in the $v_{\rm r}-v_{\rm t}$ plane corresponds to
  the observed observational and the circles represent the positions
  of the pairs from the constrained simulations. The contour line in
  the $e_{\rm tot}-l_{\rm orb}$ plane encloses $68\%$ of the Monte
  Carlo generated points to estimated the uncertainties from the
  observational values summarized in Table 1.}
\label{fig:rt}
\end{center}

@@

\begin{table*}
\caption{Summary of the comparison of the observational results
  against $\Lambda$CDM.}
\begin{center}
\begin{tabular}{ccccc}\hline
Physical & (\%) Pairs consistent & (\%) Pairs consistent & (\%) Pairs
with highest & (\%) with highest\\ 
property & with observations & with observations & likelihood in
$\Lambda$CDM & likelihood in $\Lambda$CDM\\ 
 & (full sample) & (reduced sample) & (full sample) & (reduced sample)\\ \hline
$v_{\rm r}$-$v_{\rm t}$ & (0.4\%) 8/1923 & ($<$0.6\%)0/158 & (1\%)23/1923 & (8\%)13/158\\
$e_{\rm tot}$-$l_{\rm orb}$ & (15\%)298/1923 & (24\%)38/158 & - & -\\
$\log_{10}\lambda$ & (13\%)257/1923 & (9\%)15/158 & - & -\\
$r_{\rm t}=v_{\rm t}/v_{\rm r}$& (12\%)242/1923 & (8\%)13/158 & -& -\\\hline
\end{tabular}
\end{center}
\vspace{1mm}
Note: In columns 4 and 5 the number of pairs around the preferred region
in $v_{\rm r}-v_{\rm t}$ in $\Lambda$CDM (as shown in
Fig. \ref{fig:rt} and summarized in Table \ref{table:2}) are
calculated within the same observational absolute uncertainty.
\label{table:3}
\end{table*}

@@

Figure \ref{fig:rt} summarizes the central finding of this
Letter. Most of the pairs in the two samples constructed from the
Bolshoi Simulation have radial and tangential velocities notably
different from the observational constraints.

@@

The number of pairs compatible with LG observations are listed in
Table \ref{table:3} (columns 2 and 3). In the same table (columns 4
and 5) we summarize the number of pairs around the $\Lambda$CDM values
within the same range of absolute observational uncertainty
(i.e. $\sigma_{\rm tan}=17$\kms and $\sigma_{\rm rad}=4$\kms.)

From these results we infer that the pairs around the preferred phase
space for $\Lambda$CDM are at least $13$ times more common than pairs
with the observed velocities for the Local group. We highlight that
this is a lower bound given that in the reduced sample none of the
pairs is found in the interval allowed by observations. The high
eccentric orbit of the observed LG  constitutes an unlikely
configurations for the $\Lambda$CDM LG-like objects. This holds for
the full and the reduced sample of objects.

@@

The expected observational value for $\lambda$ is lightly inconsistent with
the statistics derived from halo pairs in $\Lambda$CDM. The spin
parameter of the simulated LG-like objects is skewed towards higher
values compared with the observational estimation.  Inspection
of Figure \ref{fig:lambda} also shows that the $\lambda$ distribution is
close to a log-normal one and thus resembles the distribution of the
spin parameter of DM halos in simulation \citep{Bett2007}.

The estimated value of $\log_{10}\lambda$ of the
observed LG is $-1.72\pm 0.07$, lower than the mean value of
$\log_{10}\lambda$ of $-1.47\pm0.13$ ($-1.34\pm 0.12$) for the full
(reduced) sample. The values of the spin parameter of the 3
constrained LGs are $\log_{10}(\lambda_{\rm  CLUES})=-2.21$, $-1.72$
and $-1.65$, with the latter value belonging belonging toto the sole
constrained LG of the reduced sample.

@@

Using the same absolute values for the uncertainty in the observed
velocity components, we find that none of the halos in the reduced sample
is compatible with observations. This makes pairs with the preferred
$\Lambda$CDM values at least $13$ times more common than pairs
compatible with the observational constraint. Additionally, pairs with
a fraction of tangential to radial velocity $f_{\rm t}<0.32$ (similar
to observations) represent $8\%$ to $12\%$ of the pairs.

@@

Summarizing, we see a broad agreement in the total angular
momentum and energy and a marked difference with the precise balance
between today's radial and tangential velocities. Under the
approximation of conservation of the orbital angular momentum and
mechanical energy, this could only be explained if the initial
conditions for the formation of the Local Group are special in
comparison to the initial conditions of any other pair of dark matter
halos in the $\Lambda$CDM cosmology. This could be related to a
particular place of the Local Group in the Cosmic Web
\citep{Forero-Romero09,Hoffman12}, opening a new window into the
question of how unique, if at all, is the LG in a cosmological
context. This will continue to be studied within the CLUES framework.

\label{sec:conclusions}
\section*{Acknowledgments}  
JEF-R acknowledges financial support from the Vicerrector\'{\i}a de
Investigaciones at Universidad de los Andes through its {\it Fondo de
  Apoyo a Profesores Asistentes} and the Peter and Patricia Gruber
Foundation through its fellowship administered by the International
Astronomical Union. JEF-R also acknowledges early discussions with
Alejandro Garc\'ia that motivated this work. YH has been supported by
the ISF (1013/12)

@@

The data and source code and instructions to replicate the results of
this paper can be found as a github repository {\texttt{https://github.com/forero/LG\_Kinematics/}}. Thanks to the IPython
community \citep{IPython}. Thanks to Jessica Kirkpatrick for releasing
her Python code to make nice plots of 2D histograms.

>>>>

\begin{abstract}
Recent observations constrained the tangential velocity of M31 with
respect to the Milky Way (MW) to be $v_{\rm M31,tan}<34.4$ \kms and
the radial velocity to be in the range $v_{\rm M31,rad}=-109\pm
4.4$\kms \citep{vanderMarel12}. In this study we use a large volume
high resolution N-body cosmological simulation (Bolshoi) together with three
constrained simulations to statistically study this kinematics in the
context of the $\Lambda$CDM. The comparison of the ensembles of
simulated pairs with the observed LG {\bf at the 1-$\sigma$ level in
  the uncertainties} has been done with respect to the
observed radial and tangential velocities, the reduced orbital energy
($e_{\rm tot}$), angular momentum ($l_{\rm orb}$) and the
dimensionless spin parameter, $\lambda$. Our main results are: (i) the
preferred radial and tangential velocities for pairs in $\Lambda$CDM
are $v_{\rm r}=-80\pm 20$\kms, $v_{\rm   t}=50\pm 10$\kms, (ii) pairs
around that region are $3$ to $13$ times more common than pairs within
the observational values, (iii) $15\%$ to $24\%$ of LG-like pairs in
$\Lambda$CDM have energy and angular momentum consistent with
observations while (iv) $9\%$ to $13\%$ of pairs in the same sample
show similar values in the inferred dimensionless spin parameter. It
follows that within current observational uncertainties the
quasi-conserved quantities that characterize the orbit of
the LG, i.e. $e_{\rm tot}$, $l_{\rm   orb}$ and $\lambda$, do not
challenge the standard $\Lambda$CDM model, but the model is in tension
with regard to the actual values of the radial and tangential
velocities. This might hint to a problem of the $\Lambda$CDM model to
reproduce the observed LG.

@@

The Milky Way (MW) and Andromeda galaxy (M31) are the dominant
galaxies in the Local Group (LG). Astronomical observations of their
mass distribution impose constraints on the standard cosmological
model. The satellite overabundance problem \citep{Klypin99,Moore99},
tidal disruption features \citep{pandas09} and the disk dominated
morphology \citep{Kazantzidis2008} are examples of how LG studies
are linked to the cosmological context. Detailed studies on the
Magellanic Clouds dynamics and their possible link to M31 add to the
interest of understanding the details of the LG kinematics and
dynamics
\citep{Besla2007,Tollerud2011,Knebe2011,Fouquet2012,Teyssier2012}. However,
a general concern in the use of the LG as a tool for near-field
cosmology \citep{Freeman2002,Peebles2010} is how typical is the LG
regarding the properties of interest
\citep{Busha2011,Liu2011,ForeroRomero2011,Purcell2012}.

A new valuable piece of information in this issue is the recent
observational determination of the proper-motion measurements of M31,
which until recently had been out of reach \citep{vanderMarel12}.  The
reported measurements set an upper bound for the tangential velocity
of M31 with respect to the MW of $v_{\rm tan,M31}\leq 34.4$ km
s$^{-1}$. Together with the values of the relative radial velocity of
$v_{\rm rad,M31}=-109\pm 4.4$ \kms observations show that the relative
motion of the MW and Andromeda is consistent with a head-on
collision. With this information it is possible to quantify how common
such a kinematic configuration is in a $\Lambda$CDM Universe.

This Letter presents such a study. We use a large volume, high resolution
dark matter only N-body simulation in the concordance $\Lambda$CDM
cosmology to find a set of halo pairs with similar  characteristics as
inferred in the LG. We quantify these results in terms of the number
of pairs with given radial and tangential velocities in the
galactocentric rest frame. We also find the pairs that are consistent
with a head on collision in terms of the ratio of the radial to
tangential velocity $f_{\rm t}\equiv v_{\rm tan}/v_{\rm rad}<0.32$ and
present these results in terms of the reduced angular momentum,
mechanical energy and dimensionless spin parameter.

@@

This Letter is structured as follows. In the next section we present
the N-body simulations and the criteria we use to select LG-like halo
pairs. In Section \ref{sec:results} we present the results for the
dynamics in these pairs in terms of the tangential/radial velocities
and the orbital angular momentum/mechanical energy. In the same
section we summarize these dynamical results in terms of the
dimensionless spin parameter of the pairs. Finally, in the last
section we comment and conclude about the implications of these
results in the context of the $\Lambda$CDM model.

@@

\begin{table}
\caption{Summary of kinematic observational constraints.}
\begin{center}
\begin{tabular}{ccc}\hline
$v_{\rm M31,rad}$ &(\kms) & $-109.3\pm 4.4$\\
$v_{\rm M31,tan}$ &(\kms) & $<34.4$\\
$r_{\rm M31}$ &(kpc) & $770\pm 40$\\
${\bf r}_{\rm M31}$ & (kpc) &$(-378.9, 612.7, -283.1)$\\
$\sigma_{{\bf r}, {\rm M31}}$ & (kpc) &$(-18.9, 30.6, 14.5)$\\
${\bf v}_{\rm M31}$ & (\kms) & $(66.1, -76.3, 45.1)$\\
$\sigma_{{\bf v},{\rm M31}}$ & (\kms) &$(26.7, 19.0, 26.5)$\\
$M_{\rm 200, MW}$ & ($10^{12}\Msun$) & $1.6\pm0.5$ \\
$M_{\rm 200, M31}$ & ($10^{12}\Msun$) & $1.6\pm0.5$ \\
$M_{\rm 200,MW} + M_{\rm 200, M31}$ & ($10^{12}\Msun$) & $3.14\pm 0.58$\\
$\log_{10}\lambda$& & $-1.72\pm 0.07$ \\\hline
\end{tabular}\\
\end{center}
\vspace{1mm}
Notes:
\begin{enumerate}
\item The kinematic
  properties for M31 are reported in the galactocentric restframe 
  \citep{vanderMarel12}.

\item Values in parenthesis correspond to vector
  components. $\sigma_{\bf x}$ represents the uncertainty on the
  components of vector ${\bf x}$. {\bf The uncertainties correspond to
    1-sigma values.}  
\item The values for the individual halo
masses are consistent with the priors used by \cite{vanderMarel12}.

\item The observational uncertainties in the position vector
correspond to a $5\%$ in each component consistent with the
1-$\sigma$ uncertainties in the distance \citep[see references
  in][]{vanderMarel08}.

@@

\begin{figure*}
\begin{center}
\includegraphics[keepaspectratio=true,width=0.42\textwidth]{fig_1a.eps}
\includegraphics[keepaspectratio=true,width=0.42\textwidth]{fig_1b.eps}
\includegraphics[keepaspectratio=true,width=0.42\textwidth]{fig_1c.eps}
\includegraphics[keepaspectratio=true,width=0.42\textwidth]{fig_1d.eps}
\caption{Upper panels: 2D Histograms of the radial and tangential
  velocities for LG-like halo pairs in the Bolshoi simulation.   Lower
  panels: 2D Histograms of the orbital angular momentum ($l_{\rm orb}$)
  and mechanical energy ($e_{\rm tot}$) per unit of reduced mass
  calculated considering the halos as point masses. {\bf Left (right) panels
    correspond to the full (reduced) sample}. The bar indicates
  the number of pairs in each cell. The locii for the high
  density regions in the $v_{\rm r}-v_{\rm t}$ plane are listed in
  Table \ref{table:2}.   
  The half ellipse in the $v_{\rm r}-v_{\rm t}$ plane corresponds to
  the observations {\bf with 1-$\sigma$ uncertainties}. The contour
  line in   the $e_{\rm tot}-l_{\rm orb}$ plane encloses $68\%$ of the
  Monte   Carlo generated points to estimated the uncertainties from
  the   observational values summarized in Table 1. The circles
  represent the positions of the pairs from the constrained simulations.
}
\label{fig:rt}
\end{center}

@@

\begin{table*}
\caption{Summary of the comparison of the observational results
  against $\Lambda$CDM.}
\begin{center}
\begin{tabular}{ccccc}\hline
Physical & (\%) Pairs consistent & (\%) Pairs consistent & (\%) Pairs
with highest & (\%) with highest\\ 
property & with observations (1-$\sigma$) & with
observations (1-$\sigma$)& likelihood in $\Lambda$CDM & likelihood in $\Lambda$CDM\\ 
 & (full sample) & (reduced sample) & (full sample) & (reduced sample)\\ \hline
$v_{\rm r}$-$v_{\rm t}$ & (0.4\%) 8/1923 & ($<$0.6\%)0/158 & (1\%)23/1923 & (8\%)13/158\\
$e_{\rm tot}$-$l_{\rm orb}$ & (15\%)298/1923 & (24\%)38/158 & - & -\\
$\log_{10}\lambda$ & (13\%)257/1923 & (9\%)15/158 & - & -\\
$r_{\rm t}=v_{\rm t}/v_{\rm r}$& (12\%)242/1923 & (8\%)13/158 & -& -\\\hline
\end{tabular}
\end{center}
\vspace{1mm}
Notes:\\
\begin{enumerate}
\item 
{\bf In columns 2 and 3 consistency is defined from 1-$\sigma$
  uncerainties for $v_{\rm r}$-$v_{\rm t}$, $e_{\rm tot}$-$l_{\rm
    orb}$ and $\log_{10}\lambda$. In the case of $r_{\rm t}$ is
  defined from the constrain $r_{\rm t}<0.32$ derived from the
  1$-\sigma$ observational uncertainties in the radial and tangential
  velocities.} 
\item
In columns 4 and 5 the number of pairs around the preferred region
in $v_{\rm r}-v_{\rm t}$ in $\Lambda$CDM (as shown in
Fig. \ref{fig:rt} and summarized in Table \ref{table:2}) are
calculated within the same observational 1-$\sigma$ absolute uncertainty.
\end{enumerate}
\label{table:3}
\end{table*}

%Obs (full -3sigma) 127
%Theory (full 3sigma) 130
%Obs (reduced -3sigma) 11
%Theory (reduced 3sigma) 17

@@

Figure \ref{fig:rt} summarizes the central finding of this
Letter. Most of the pairs in the two samples constructed from the
Bolshoi Simulation have radial and tangential velocities notably
different from the observational constraints at the 1-$\sigma$ level.

@@

The number of pairs compatible with LG observations {\bf at the
  1-$\sigma$ level} are listed in
Table \ref{table:3} (columns 2 and 3). In the same table (columns 4
and 5) we summarize the number of pairs around the $\Lambda$CDM values
within the same range of absolute observational uncertainty 
(i.e. $\sigma_{\rm tan}=17$\kms and $\sigma_{\rm rad}=4$\kms.)

From these results we infer that the pairs around the preferred phase
space for $\Lambda$CDM are at least $13$ times more common than pairs
with the observed velocities for the Local group. We highlight that
this is a lower bound given that in the reduced sample none of the
pairs is found in the interval allowed by observations. The high
eccentric orbit of the observed LG  constitutes an unlikely
configurations for the $\Lambda$CDM LG-like objects. This holds for
the full and the reduced sample of objects. {\bf This
  conclusions are valid at 1-$\sigma$ level in the observational
  uncertainties. Taking 3-$\sigma$ levels (i.e. uncertainties on the
  order of $\sigma_{\rm tan}=51$\kms and $\sigma_{\rm
    rad}=12$\kms the observational values are as likely as the
  preferred $\Lambda$CDM values, although only $\sim 6\%$ of the pairs
in the full and reduced samples would be consistent with observations.}

@@

The expected observational value for $\lambda$ is slightly inconsistent with
the statistics derived from halo pairs in $\Lambda$CDM. The spin
parameter of the simulated LG-like objects is skewed towards higher
values compared with the observational estimation.  Inspection
of Figure \ref{fig:lambda} also shows that the $\lambda$ distribution is
close to a log-normal one and thus resembles the distribution of the
spin parameter of DM halos in simulation \citep{Bett2007}.

The estimated value of $\log_{10}\lambda$ of the
observed LG is $-1.72\pm 0.07$, lower than the mean value of
$\log_{10}\lambda$ of $-1.47\pm0.13$ ($-1.34\pm 0.12$) for the full
(reduced) sample. The values of the spin parameter of the 3
constrained LGs are $\log_{10}(\lambda_{\rm  CLUES})=-2.21$, $-1.72$
and $-1.65$, with the latter value belonging to the sole
constrained LG of the reduced sample.

@@

Using the same absolute values for the uncertainty in the observed
velocity components, we find that none of the halos in the reduced sample
are compatible with observations. This makes pairs with the preferred
$\Lambda$CDM values at least $13$ times more common than pairs
compatible with the observational constraint. Additionally, pairs with
a fraction of tangential to radial velocity $f_{\rm t}<0.32$ (similar
to observations) represent $8\%$ to $12\%$ of the pairs.

@@

Summarizing, we see a broad agreement in the total angular
momentum and energy and a marked difference with the precise balance
between today's radial and tangential velocities, {\bf i.e. head-on
  collisions of MW-like halos are not common in $\Lambda$CDM}.

Under the
approximation of conservation of the orbital angular momentum and
mechanical energy, this could only be explained if the initial
conditions for the formation of the Local Group are special in
comparison to the initial conditions of any other pair of dark matter
halos in the $\Lambda$CDM cosmology. {\bf A thorough characterization
  of the small pair sample consistent with observations in terms
  of halo properties such as concentration, spin and place in the cosmic web
  \citep{Forero-Romero09,Hoffman12} is  underway (Bustamante et al. in
  prep.)}.

{\bf The results presented in this Letter open a new window into the
  question of how unique, if at all, is the LG in a cosmological
  context. This will continue to be studied within the CLUES project
  framework.}

\label{sec:conclusions}
\section*{Acknowledgments}  
JEF-R acknowledges financial support from Universidad de los Andes through its {\it Fondo de
  Apoyo a Profesores Asistentes} and the Peter and Patricia Gruber
Foundation through its fellowship administered by the IAU. JEF-R also acknowledges early discussions with
Alejandro Garc\'ia that motivated this work. YH has been supported by
the ISF (1013/12)

@@

The data and source code and instructions to replicate the results of
this paper can be found here {\texttt{https://github.com/forero/LG\_Kinematics/}}. Thanks to the IPython
community \citep{IPython}. Thanks to Jessica Kirkpatrick for releasing
her Python code to make nice plots of 2D histograms.
